---
title: "Quakinator - your handy earthquake predictor"
author: "Christian Karcher"
font-family: 'Helvetica'
date: "14 November 2017"
output: slidy_presentation
footer: "C. Karcher, Nov. 2017"
css: styles.css
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Ever wondered how much the earth is going to shake next year? {.slide2 }

No? Well, you should! And good news, everybody, now there's a tool for that: <https://chrisokay.shinyapps.io/quakinator/>.

And not only that, you will be able to

- filter by Richter scale!
- filter by year!
- and (you guessed it) filter by **tsunamis**!


## Filters you say?! {.slide3 }

Yes, friend, you heard right! The compelling language of R is able to filter!
Let me demonstrate: By carefully reading the data, one is able to select certain characteristic data from it!

```{r echo = TRUE}
earthquakes <- read.csv(file='../shinyEarthquake.csv', sep = '\t')
dim(earthquakes)
tsuFilter = 'Tsu'
filtered <- earthquakes[earthquakes$FLAG_TSUNAMI %in% tsuFilter,]
dim(filtered)
```
Unbelievable, right?

## But what about the data? {.slide4 }

The data has been carefully collected by reliable sources all over the planet. Either that, or I meanwhile have forgotten the original source, must have been taken from somewhere here: <https://earthquake.usgs.gov/data/data.php>

I stripped most of the data without any Richter scale information and ended up with a dataset of 7000+ entries, which was suffice for this little project. Originally the data contained information about deaths and damages (which reminded me an awful lot of this "tornado project" from the cleaning data course), but I decided not to make things too complicated and stripped it down to "richter and tsunamis" only.

## Revealing the Backend {.slide5 }

To get insight into the whole application, the sourcecode has been made public here: <https://github.com/ChrisOKay/coursera_dp_project>

Additionally, here's the main calculation of server.R, which fits a linear model to the filtered data (x=year, y=Richter scale (EQ_PRIMARY)):

```{r echo = TRUE, eval = FALSE}
linFit <- reactive({
  data <- aggregate(EQ_PRIMARY ~ YEAR, filteredData(), mean)
  lm(EQ_PRIMARY ~ YEAR, data)
})
output$mypredict <- renderText({ 
  round(predict(linFit(),data.frame(YEAR=2018))*10)/10
})
```

*Thank you* for your attention and have a great time with the tool!


